Enabling conditions for effective community based marine reserves in small scale fisheries

### 2018-02-14 Frontiers_Manuscript/Villasenor-Derbez.tex
ÑAdded first final version of manuscript draft

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,*}\), Stuart Fulton\(^{2}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\title[Mexican marine reserves]{Cool title here}

@@

Abstract here

@@

Marine reserve science has largely focused on understanding the
ecological effects of these areas, which include increased biomass,
richness, and densities of organisms within the protected regions
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change
mitigation \citep{roberts_2017-J9}, and protection from environmental
variability \citep{micheli_2012-EU}. However, there is considerably less
literature focusing on the relationship between socioeconomic and
governance structures and their relationship to ecological
effectriveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or
benefits to fisheries \citep{krueck_2017-J1}; evaluations of marine
reserves rarely provide a hollistic view of the social-ecological system
\citep{lpezangarita_2014}. Here, we combine causal inference techniques
\citep{depalma_2018} and the social-ecological systems framework
\citep{ostrom_2009-hg} to provide a comprehensive ecological and
socioeconomic evaluation of four community-based marine reserves in
three coastal communities in Mexico.

Marine Reserves in Mexico have been commonly implemented as ``core
zones'' within Biosphere Reserves (BRs) that are administered by the
National Commision of Protected Areas (\emph{Comisión Nacional de Áreas
Marinas Protegidas}, CONANP). While CONANP has made efforts to have a
participatory process, the implementation of these zones is still
characterized by top--down approaches. This motivated Civil Society
Organizations (CSOs) to work with coastal communities to implement
community--based marine reserves \citep{uribe_2010-u2}, which are
usually established within a Territorial Use Rights for Fisheries
(TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
This bottom--up approach allows fishers to disign their own reserves,
which increases compliance and self--enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
these reserves still lack legal recognition, making them vulnerable to
poaching. In 2014, a new norm \citep{nom} allowed fishers to request the
legal recognition of a community--based reserve under the name of
``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}, FR). This new norm
thus combines bottom--up approaches to design marine reserves, along
with a legal recognition of the management intervention. Since then,
\emph{39} FR have been implemented along the Pacific, Gulfof
Caloifornia, and Mexican Caribbean coastlines, but no formal evaluation
of their effectiveness has taken place.

While there are ecological factors defining the success of a MR
(\emph{i.e.} habitat representation, initial state of protection,
connectivity to other protected areas), their effectivness also depends
on the socioeconomic and governance settings under which they are
implemented. Literature shows that many non-ecological characteristics
can play an equally important role in the effectivenes of MRs. For
example, age of a reserve (\emph{i.e.} time since its implementation),
size, and habitat contained were key to the effectiveness of MRs in
Palau \citep{friedlander_2017-oI}. In the Mediterranean,
\citet{difranco_2016-Xw} identify that surveilance and enfocrement,
presence of a management plan, and involvement of fishers in management
and decision--making along with promotion of sustainable fishing
practices were the key fators that increased stock health and income to
fishers. At a global level, \citet{edgar_2014-UO} indicate that
enforcement, age, size, and isolation were important factors determining
effectiveness of the reserves.

The objective of this work is twofold: i) Provide the first evaluation
of community--based marine reserves in Mexico, and ii) provide a
comprehensive evaluation of the social--ecological system to identify
how sociocoenomic and governance characteristics relate to ecological
effectiveness. With the purpose of providing a hollistic evaluation, we
combine ecological, socioeconomic, and governance indicators. We use
causal inference techniques to provide a measurement of the effect of
the management intervention, and combine it with the social--ecological
systems framework \citep{ostrom_2009-hg}.

@@

We focus our evaluation in three coastal communities from the Pacific
coast of Baja California (n = 1) and the Mesoamerican Reef System (n =
2; Fig \ref{fig:map}). Isla Natividad (IN) lies west of the Baja
California Peninsula (Fig \ref{fig:map}B), where kelp forests
(\emph{Macrocystis pyrifera}) and rocky reefs are the predominant
habitats. The island is home to a fishing cooperative (\emph{Sociedad
Cooperative de Producción Pesquera Buzos y Pescadores de la Baja
California SCL}), that holds a TURF for spiny lobster (\emph{Panulirus
interruptus}). However, other resources like finfish (yellowtail jack,
\emph{Seriola lalandi}), sea cucumber (\emph{Parastichopus
parvimensis}), red sea urchin (\emph{Mesocentrotus franciscanus}), snail
(\emph{Megastraea turbanica} y \emph{M. undosa}), and abalone
(\emph{Haliotis spp}, until 2010) are also important sources of income.
In 2006, the community decided to implement two community--based marine
reserves within their fishing grounds, seeking to recover depleted
stocks of invertebrate species (mainly lobster and abalone). Until
today, these reserves are yet to be legally recognized as Fishing
Refugia.

The other two communities are Maria Elena (ME; Fig \ref{fig:map}C) and
Punta Herrero (PH; Fig \ref{fig:map}D) in the Yucatan Peninsula, where
coral reefs and mangroves are the representative coastal ecosystems. ME
is a fishing camp --visited intermitently during the fishing season--
belonging to the Cozumel fishing cooperative. PH is home to the ``José
María Azcorra'' fishing cooperative. The main source of income to both
communities is the caribbean spinly lobster fishery (\emph{Panulirus
argus}), which is carried out within their respective TURFs. These
communities also target finfish in the off season, mainly snappers
(Lutjanidae) and groupers (Serranidae). ME established eight marine
reserves in 2012, and PH established four marine reserves in 2013. All
these reserves are legally recognized as Fishing Refugia.

@@

To perform the evaluation of these reserves we use thre sources of
information. Ecological data come from the annual ecological monitorings
of reserve and control areas, carried out by members from each community
and personnel from the Mexican CSO ``Comunidad y Biodiversidad''
(\href{www.cobi.org.mx}{COBI}). These monitorings record richness and
abundances of fish and invertebrate species in the reservs and control
sites. For fish census, size structures are also collected to derive
biomass. We define control sites as regions with habitat characteristics
similar to the corresponding reserves, and that presumably had the same
probability of being selected as reserves during the design phase. From
all the reserves in these three communities, we use the ones that have
data for reserve and control sites before and after the implementation
of the reserve. This provides us with a Before--After--Control--Impact
(\emph{i.e.} BACI) design that allows us to capture and control for
temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI
designs and causal inference techniques have proven effective to
evaluate marine reserves, as they allow us to causaly attribute observed
changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All reserves were
surveyed annualy from at least one year before implementation until
2016. Table \ref{tab:com_sum} shows a summary of the number of reserves,
year of implementation, and number of transects for each reserve.

@@

\caption{\label{tab:unnamed-chunk-2}\label{tab:com_sum} Summary of commuity--based marine reserves by community. Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.}
\centering
\begin{tabular}[t]{l|l|r|r|r|r|r|r}
\hline
Community & Reserve - Control & Imp & Start & Cf & Rf & Ci & Ri\\
\hline
Isla Natividad & La Plana / Las Cuevas - La Dulce / Babencho & 2006 & 2006 & 405 & 242 & 415 & 245\\
\hline
Maria Elena & Cabezo - Cabezo (Control) & 2012 & 2012 & 44 & 45 & 27 & 21\\
\hline
Punta Herrero & El Faro - El Faro (Control) & 2013 & 2013 & 39 & 40 & 24 & 32\\
\hline
Punta Herrero & Manchon - Manchon (Control) & 2013 & 2012 & 43 & 45 & 27 & 42\\
\hline
\end{tabular}
\end{table}

@@

\textbf{Governance data were collected at the community--level, using
key informants to collect the necesary information on X, Y, and Z. Algo
mas de Ere.}

@@

Following a framework that relates reserve objectives to performance
indicators \citep{Villasenor-Derbez_2018}, we use five biological, two
socioeconomic, and five governance indicators to evaluate these marine
reserves Table \ref{table:indicators}.

@@

\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l}
\hline
Category & Indicador\\
\hline
Biological & Abundance\\
\hline
 & Richness\\
\hline
 & Shannon's diversity index\\
\hline
 & Biomass\\
\hline
 & Abundance of target species (lobsters)\\
\hline
Socioeconomic & Income from target species\\
\hline
 & Landings from target species\\
\hline
Governance & Type of access to the fishery\\
\hline
 & Perceived degree of illegal fishing\\
\hline
 & Reserve surveilance anf enforcement\\
\hline
 & Type of fishing organization\\
\hline
 & Age of the reserve\\
\hline
\end{tabular}
\end{table}

@@

Where year--fixed effects are represented by \(\gamma_{it} Year_t\), and
\(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\))
and control (\(Zone = 0\)) sites. The interaction term
\(\lambda_{it} Year_t\times Zone_i\) represents represent the mean
change in the indicator inside the reserve, for year \(t\), with respect
to the first year of evaluation in the control site (See Table
\ref{tab:com_sum}). When evaluating biomass and abundances, we include
species--fixed effects (\(\sigma_j\)). For abundances and richness
(\emph{i.e.} count data) the model is estimated with a quasipoisson
error distribution.

@@

\subsubsection{Governance}\label{governance}

\textbf{Texto de ere y los SES}

@@

Our methodological approach with biological indicators allows us to make
a causal link between the implementation of marine reserves and the
observed trends by accounting for temporal and spatial dynamics
\citep{depalma_2018}. The effect of the reserve is captured by the
\(\lambda_t\) coefficient, and represents the difference observed
between the control site before the implementation of the reserve and
the reserve site at time \(t\) after controling for other time and space
variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively). Here
we present the effect that marine reserves had on each of the biological
indicators for each coastal community, along with the trends in
socioeconomic indicators of lobster catches and revenues. We also
provide an overview of the state of the socioeconomic and governance
settings of each community, and discuss how these dimensions might be
intertwined with each other.

@@

Effect sizes for biological indicators are shown in Figure
\ref{fig:indicators}, and Figure \ref{fig:res_com} shows the summarized
biological effects by community. Isla Natividad shows inconsistent
effects across data sources (\emph{i.e.} fish vs.~invertebrates). For
example, the reserve had a small effect on fish abundances (Fig
\ref{fig:indicators}A), where only year 2010 showed significat effect
sizes in fish abundances (\(p<0.05\)) and all other years oscillated
above and under zero (\(p > 0.05\)). However, invertebrate abundances
(Fig \ref{fig:indicators}B) presented a positive trend relative to the
control site before implementation (\(p < 0.05\)) for all but 1 year
(2008). Maria Elena and Punta Herrero showed no significant increase in
fish and invertebrate abundances (\(p< 0.05\)), except for invertebrates
in Punta Herrero for 2014 --right after the implementation of the
reserves-- which showed a significant increase (\emph{i.e.}
\(\lambda_{2014} = 2.5\), \(p < 0.05\)). Full tables with model
coefficients are presented in the supplementary materials (\textbf{S1
Table}, \textbf{S2 Table}, \textbf{S3 Table}).

@@

Biomass was only evaluated for fish data (Fig \ref{fig:indicators}G). In
Isla Natividad, fish biomass presented a steady but small increase
(\(p>0.05\)), and exhibited an increased variability in biomass between
2013 and 2016. Maria Elena and Punta Herrero also showed small,
non-statistically significant increases in fish biomass (\(p>0.05\)).
The last biological indicator is abundance of target species,
\emph{Panulirus interruptus} and \emph{P. argus}, for the Pacific and
Caribbean, respectively (Fig \ref{fig:indicators}H). Isla Natividad
presented small constantly-positive effects but were not significantly
different from the reference point of control site before the
implementation of the reserve (\(p > 0.05\). Maria Elena showed
significant increases in lobster densities in the order of
\(\lambda_t = 10\) (\(p < 0.05\)). Finally, punta herrero presented
alternating negative and positive effects, but these were not different
from the baseline case (\(p > 0.05\)).

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}\label{fig:indicators}Effect sizes for
marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
blue triangles), and Punta Herrero (PH; green squares) for
community-levell indicators. Plots are ordered by survey type (left:
fish; right: invertebrates) and indicators: Abundance (A, B), Richness
(C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster
(\emph{Panulirus spp}) abundances (H). Points are jittered hotizontally
to avoid overplotting. Points indicate the effect size, and errorbars
are heteroskedastic-robust standard errors.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}\label{fig:res_com}Summarized effects of
the marne reserves by direction (positive - negative) and significance.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}\label{fig:lobsters}Time series of
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green
squares).}
\end{figure}

\subsection{Governance}\label{governance-1}

>>>>

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,*}\), Álvin Suarez-Castillo\(^{2}\), Stuart
Fulton\(^{2}\), Jorge Torre\(^{2}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

\title[Marine reserve effectiveness]{Enabling conditions for effective community based marine reserves in
small scale fisheries}

@@

Abstract after comments from co-authors

@@

Marine reserve science has largely focused on understanding the
ecological effects of these areas, which include increased biomass,
richness, and densities of organisms within the protected regions
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, climate change
mitigation \citep{roberts_2017-J9}, and protection from environmental
variability \citep{micheli_2012-EU}. However, there is considerably less
literature focusing on the relationship between socioeconomic and
governance structures and their relationship to ecological effectiveness
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_} or benefits to
fisheries \citep{krueck_2017-J1}; evaluations of marine reserves rarely
provide a holistic view of the social-ecological system
\citep{lpezangarita_2014}. Here, we combine causal inference techniques
\citep{depalma_2018} and the social-ecological systems framework
\citep{ostrom_2009-hg} to provide a comprehensive ecological and
socioeconomic evaluation of four community-based marine reserves in
three coastal communities in Mexico.

Marine Reserves in Mexico have been commonly implemented as ``core
zones'' within Biosphere Reserves that are administered by the National
Commission of Protected Areas (\emph{Comisión Nacional de Áreas Marinas
Protegidas}, CONANP). While CONANP has made efforts to have a
participatory process, the implementation of these zones is still
characterized by top--down approaches. This motivated Civil Society
Organizations (CSOs) to work with coastal communities to implement
community--based marine reserves \citep{uribe_2010-u2}, which are
usually established within a Territorial Use Rights for Fisheries
(TURFs); thus making them TURF--reserves \citep{afflerbach_2014-HP}.
This bottom--up approach allows fishers to design their own reserves,
which increases compliance and self--enforcement
\citep{gelcich_2015-Gw,espinosaromero_2014-PY,beger_2004-Y8}. However,
these reserves still lack legal recognition, making them vulnerable to
poaching. In 2014, a new norm \citep{nom} allowed fishers to request the
legal recognition of a community--based reserve under the name of
``Fishing Refugia'' (\emph{Zona de Refugio Pesquero}, FR). This new norm
thus combines bottom--up approaches to design marine reserves, along
with a legal recognition of the management intervention. Since then,
\emph{39} FR have been implemented along the Pacific, Gulf of
California, and Mexican Caribbean coastlines, but no formal evaluation
of their effectiveness has taken place.

While there are ecological factors defining the success of a MR
(\emph{i.e.} habitat representation, initial state of protection,
connectivity to other protected areas), their effectiveness also depends
on the socioeconomic and governance settings under which they are
implemented. Literature shows that many non-ecological characteristics
can play an equally important role in the effectiveness of MRs. For
example, age of a reserve (\emph{i.e.} time since its implementation),
size, and habitat contained were key to the effectiveness of MRs in
Palau \citep{friedlander_2017-oI}. In the Mediterranean,
\citet{difranco_2016-Xw} identify that surveillance and enforcement,
presence of a management plan, and involvement of fishers in management
and decision--making along with promotion of sustainable fishing
practices were the key factors that increased stock health and income to
fishers. At a global level, \citet{edgar_2014-UO} indicate that
enforcement, age, size, and isolation were important factors determining
effectiveness of the reserves.

The objective of this work is twofold: i) Provide the first evaluation
of community--based marine reserves in Mexico, and ii) provide a
comprehensive evaluation of the social--ecological system to identify
how socioeconomic and governance characteristics relate to ecological
effectiveness. With the purpose of providing a holistic evaluation, we
combine ecological, socioeconomic, and governance indicators. We use
causal inference techniques to provide a measurement of the effect of
the management intervention, and combine it with the social--ecological
systems framework \citep{ostrom_2009-hg}.

@@

We focus our evaluation in three coastal communities from the Pacific
coast of Baja California (n = 1) and the Mesoamerican Reef System (n =
2; Fig \ref{fig:map}). Isla Natividad (IN) lies west of the Baja
California Peninsula (Fig \ref{fig:map}B), where kelp forests
(\emph{Macrocystis pyrifera}) and rocky reefs are the predominant
habitats. The island is home to a fishing cooperative (\emph{Sociedad
Cooperative de Producción Pesquera Buzos y Pescadores de la Baja
California SCL}), that holds a TURF for spiny lobster (\emph{Panulirus
interruptus}). However, other resources like finfish (yellow-tail jack,
\emph{Seriola lalandi}), sea cucumber (\emph{Parastichopus
parvimensis}), red sea urchin (\emph{Mesocentrotus franciscanus}), snail
(\emph{Megastraea turbanica} y \emph{M. undosa}), and abalone
(\emph{Haliotis spp}, until 2010) are also important sources of income.
In 2006, the community decided to implement two community--based marine
reserves within their fishing grounds, seeking to recover depleted
stocks of invertebrate species (mainly lobster and abalone). Until
today, these reserves are yet to be legally recognized as Fishing
Refugia.

The other two communities are Maria Elena (ME; Fig \ref{fig:map}C) and
Punta Herrero (PH; Fig \ref{fig:map}D) in the Yucatan Peninsula, where
coral reefs and mangroves are the representative coastal ecosystems. ME
is a fishing camp --visited intermittently during the fishing season--
belonging to the Cozumel fishing cooperative. PH is home to the ``José
María Azcorra'' fishing cooperative. The main source of income to both
communities is the Caribbean spiny lobster fishery (\emph{Panulirus
argus}), which is carried out within their respective TURFs. These
communities also target finfish in the off season, mainly snappers
(Lutjanidae) and groupers (Serranidae). ME established eight marine
reserves in 2012, and PH established four marine reserves in 2013. All
these reserves are legally recognized as Fishing Refugia.

@@

To perform the evaluation of these reserves we use thee sources of
information. Ecological data come from the annual ecological monitoring
of reserve and control areas, carried out by members from each community
and personnel from the Mexican CSO ``Comunidad y Biodiversidad''
(\href{www.cobi.org.mx}{COBI}). These monitorings record richness and
abundances of fish and invertebrate species in the reserves and control
sites. For fish census, size structures are also collected to derive
biomass. We define control sites as regions with habitat characteristics
similar to the corresponding reserves, and that presumably had the same
probability of being selected as reserves during the design phase. From
all the reserves in these three communities, we use the ones that have
data for reserve and control sites before and after the implementation
of the reserve. This provides us with a Before-After-Control-Impact
(\emph{i.e.} BACI) design that allows us to capture and control for
temporal and spatial dynamics \citep{depalma_2018,ferraro_2006-oW}. BACI
designs and causal inference techniques have proven effective to
evaluate marine reserves, as they allow us to causally attribute
observed changes to the intervention
\citep{moland_2013-VP,Villasenor-Derbez_2018}. All reserves were
surveyed annually from at least one year before implementation until
2016. Table \ref{table:com_sum} shows a summary of the number of
reserves, year of implementation, and number of transects for each
reserve.

@@

\caption{\label{tab:unnamed-chunk-2}\label{table:com_sum} Summary of commuity--based marine reserves by community. Imp = Year of implementation, Start = Year of first sampling, number of fish transects in control (Cf) and reserve (Rf) sites, and number of invertebrate transects in Control (Ci) and Reserve (Ri) sites.}
\centering
\begin{tabular}[t]{l|l|r|r|r|r|r|r}
\hline
Community & Reserve - Control & Imp & Start & Cf & Rf & Ci & Ri\\
\hline
Isla Natividad & La Plana / Las Cuevas - La Dulce / Babencho & 2006 & 2006 & 405 & 242 & 415 & 245\\
\hline
Maria Elena & Cabezo - Cabezo (Control) & 2012 & 2012 & 44 & 45 & 27 & 21\\
\hline
Punta Herrero & El Faro - El Faro (Control) & 2013 & 2013 & 39 & 40 & 24 & 32\\
\hline
Punta Herrero & Manchon - Manchon (Control) & 2013 & 2012 & 43 & 45 & 27 & 42\\
\hline
\end{tabular}
\end{table}

@@

Governance data were collected at the community-level. The information
was compiled by combining key informants and the authors; experience and
knowledge of the communities to collect the necessary information. These
data contain information on the ecological system where the fishing
activities develop, as well as the governance structures present in the
cooperative. We also gathered information on the resource unit
(\emph{i.e.} lobsters) and the relevant actors present in each community
\citep{leslie_2015-na}.

@@

Following a framework that relates reserve objectives to performance
indicators \citep{Villasenor-Derbez_2018}, we use five biological and
two socioeconomic indicators to evaluate these marine reserves Table
\ref{table:indicators}. We also use a set of governance indicators to
analyze the governance structures of each cooperative
\citep{leslie_2015-na}. The indicators (Table \ref{table:gov_ind}) focus
on the resource system (four indicators), governance system (seven
indicators), resource units (three indicators) and actors (three
indicators).

@@

\caption{\label{tab:unnamed-chunk-3}\label{table:indicators}List of indicators used to evaluate the effectiveness of marine reserves, grouped by category.}
\centering
\begin{tabular}[t]{l|l}
\hline
Category & Indicador\\
\hline
Biological & Abundance\\
\hline
 & Richness\\
\hline
 & Shannon's diversity index\\
\hline
 & Biomass\\
\hline
 & Abundance of target species (lobsters)\\
\hline
Socioeconomic & Income from target species\\
\hline
 & Landings from target species\\
\hline
\end{tabular}
\end{table}

\caption{\label{tab:unnamed-chunk-4}\label{table:gov_ind}Indicators used for the operationalization of the SES framework \citep{leslie_2015-na}}
\centering
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{8em}|>{\raggedright\arraybackslash}p{8em}|>{\raggedright\arraybackslash}p{8em}}
\hline
Indicator & Isla Natividad & Maria Elena & Punta Herrero\\
\hline
\textbf{Resource systems (RS)} & \textbf{} & \textbf{}\\
\hline
TURF presence & Yes & Yes & Yes\\
\hline
Type of ecosystem & Kelp Forest / Rocky Reefs & Coral Reef & Coral Reef\\
\hline
Intensity of environmental Disturbance & El nino event & Hurricanes & Hurricanes\\
\hline
Location & Island & Coastal & Coastal\\
\hline
\textbf{Governance systems (GS)} & \textbf{} & \textbf{}\\
\hline
Fishing cooperative & Yes & Yes & Yes\\
\hline
Involved actors & COBI, Stanford, REBIVI & Alianza Kanan Kay, COBI, CONANP, Coop, CONAPESCA, Oceanus, FCyRH, FHMM, & Alianza Kanan Kay, COBI, CONANP, Coop, CONAPESCA, Oceanus, FCyRH, FHMM,\\
\hline
Presence of an inter-cooperative structure & Fedecoop & Non & Non\\
\hline
Fishing Regulations & Size limits, seasonal closures, quotas & Size limits, seasonal closures & Size limits, seasonal closures\\
\hline
Enforcement technology & Boats & Boats & Land enforcement\\
\hline
MR enforcement &  &  & \\
\hline
Cooperative regulations &  &  & \\
\hline
\textbf{Resource Units (RU)} & \textbf{} & \textbf{}\\
\hline
Adult targeted species mobility & 1km & 30km & 30km\\
\hline
Targeted species longevity (years) &  &  & \\
\hline
Price of targeted species &  &  & \\
\hline
\textbf{Actors (A)} & \textbf{} & \textbf{}\\
\hline
Leadership &  &  & \\
\hline
Level of illegal fishing & 1 & 1 & 3\\
\hline
Presence of alternative livelihoods &  &  & \\
\hline
\end{tabular}
\end{table}

@@

Where year--fixed effects are represented by \(\gamma_{it} Year_t\), and
\(\beta Zone_i\) captures the difference between reserve (\(Zone = 1\))
and control (\(Zone = 0\)) sites. The interaction term
\(\lambda_{it} Year_t\times Zone_i\) represents represent the mean
change in the indicator inside the reserve, for year \(t\), with respect
to the first year of evaluation in the control site (See Table
\ref{table:com_sum}). When evaluating biomass and abundances, we include
species--fixed effects (\(\sigma_j\)). For abundances and richness
(\emph{i.e.} count data) the model is estimated with a quasipoisson
error distribution.

@@

@@

Our methodological approach with biological indicators allows us to make
a causal link between the implementation of marine reserves and the
observed trends by accounting for temporal and spatial dynamics
\citep{depalma_2018}. The effect of the reserve is captured by the
\(\lambda_t\) coefficient, and represents the difference observed
between the control site before the implementation of the reserve and
the reserve site at time \(t\) after controlling for other time and
space variations (\emph{i.e.} \(\gamma_t\) and \(\beta\) respectively).
Here we present the effect that marine reserves had on each of the
biological indicators for each coastal community, along with the trends
in socioeconomic indicators of lobster catches and revenues. We also
provide an overview of the state of the socioeconomic and governance
settings of each community, and discuss how these dimensions might be
intertwined with each other.

@@

Effect sizes for biological indicators are shown in Figure
\ref{fig:indicators}, and Figure \ref{fig:res_com} shows the summarized
biological effects by community. Isla Natividad shows inconsistent
effects across data sources (\emph{i.e.} fish vs.~invertebrates). For
example, the reserve had a small effect on fish abundances (Fig
\ref{fig:indicators}A), where only year 2010 showed significant effect
sizes in fish abundances (\(p<0.05\)) and all other years oscillated
above and under zero (\(p > 0.05\)). However, invertebrate abundances
(Fig \ref{fig:indicators}B) presented a positive trend relative to the
control site before implementation (\(p < 0.05\)) for all but 1 year
(2008). Maria Elena and Punta Herrero showed no significant increase in
fish and invertebrate abundances (\(p< 0.05\)), except for invertebrates
in Punta Herrero for 2014 --right after the implementation of the
reserves-- which showed a significant increase (\emph{i.e.}
\(\lambda_{2014} = 2.5\), \(p < 0.05\)). Full tables with model
coefficients are presented in the supplementary materials (\textbf{S1
Table}, \textbf{S2 Table}, \textbf{S3 Table}).

@@

Biomass was only evaluated for fish data (Fig \ref{fig:indicators}G). In
Isla Natividad, fish biomass presented a steady but small increase
(\(p>0.05\)), and exhibited an increased variability in biomass between
2013 and 2016. Maria Elena and Punta Herrero also showed small,
non-statistically significant increases in fish biomass (\(p>0.05\)).
The last biological indicator is abundance of target species,
\emph{Panulirus interruptus} and \emph{P. argus}, for the Pacific and
Caribbean, respectively (Fig \ref{fig:indicators}H). Isla Natividad
presented small constantly-positive effects but were not significantly
different from the reference point of control site before the
implementation of the reserve (\(p > 0.05\). Maria Elena showed
significant increases in lobster densities in the order of
\(\lambda_t = 10\) (\(p < 0.05\)). Finally, Punta Herrero presented
alternating negative and positive effects, but these were not different
from the baseline case (\(p > 0.05\)).

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}\label{fig:indicators}Effect sizes for
marine reserves from Isla Natividad (IN; red cirlcles), Maria Elena (ME;
blue triangles), and Punta Herrero (PH; green squares) for
community-levell indicators. Plots are ordered by survey type (left:
fish; right: invertebrates) and indicators: Abundance (A, B), Richness
(C, D), Shannon's diversity index (E, F),fish biomass (G), and lobster
(\emph{Panulirus spp}) abundances (H). Points are jittered hotizontally
to avoid overplotting. Points indicate the effect size, and errorbars
are heteroskedastic-robust standard errors.}
\end{figure}

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}\label{fig:res_com}Summarized effects of
the marne reserves by direction (positive - negative) and significance.}
\end{figure}

@@

Lobster catches and income showed a increase after the implementation of
the reserves for Isla Natividad and Maria Elena, but not for Punta
Herrero (Fig \ref{fig:lobsters}). However, the differences in catches
and and revenue were not different in the periods before and after the
implementation (\(p > 0.05\)) except for revenues in Isla Natividad,
which presented a significant increase of 14.37 (M MXP; \(p<0.05\)).
Table \ref{table:soc_res} presents all the regression coefficients.

\begin{table}[!htbp] \centering 
  \caption{\label{table:soc_res} Regression coefficients for lobster catches and revenues for Isla Natividad (1, 4), Maria Elena (2, 5), and Punta Herrero (3, 6).} 
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{\textit{Dependent variable:}} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{3}{c}{Catches (tones)} & \multicolumn{3}{c}{Revenue(M MXP)} \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 Post & 7.37 & 5.83 & $-$1.26 & 14.37$^{***}$ & 1.24 & $-$0.06 \\ 
  Constant & 122.68$^{***}$ & 11.41$^{***}$ & 6.06$^{***}$ & 10.89$^{***}$ & 2.04$^{***}$ & 1.06$^{***}$ \\ 
 \hline \\[-1.8ex] 
Observations & 15 & 14 & 13 & 15 & 14 & 13 \\ 
R$^{2}$ & 0.02 & 0.04 & 0.01 & 0.52 & 0.04 & 0.0004 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}

\begin{figure}
\centering
\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{\label{fig:unnamed-chunk-8}\label{fig:lobsters}Time series of
lobster catches (A) and revenues (B) in at Isla Natividad (IN; red
circles), Maria Elena (ME; blue triangles), and Punta Herrero (PH; green
squares).}
\end{figure}

\subsection{Governance}\label{governance}

Although we have little information on the social dimension of these
fisheries, using the SES framework indicators (Table
\ref{table:gov_ind}), we can analyze the performance of each governance
system with respect to MR enforcement (Table \ref{table:gov_res}). In
general, the presence and success of conservation initiatives depends on
the incentives of local communities to maintain a healthy status of the
resources they depend upon \citep{jupiter_2017}. The enabling conditions
for conservation seem to be strongly present in Isla Natividad. Due to
the clarity of access rights and isolation, the benefits of conservation
directly benefit the members of the fishing cooperative. These
conditions have favored the development of an efficient community based
enforcement system. In contrast, the communities of Maria Elena and
Punta Herrero are located near other fishing communities and cities. In
Maria Elena, the fishing pressure caused by outsiders can be reduced by
implementing a strong enforcement system (in water and land) supported
by CSOs and the local government (CONANP). Lastly, the community of
Punta Herrero shows the highest levels of illegal activities which can
be attributed to its connectedness to other communities and the lack of
appropriate technologies for enforcement.
